Topological Equivariant Artist Model

### 2025-06-13 paper/main.tex
other misc fixes + abstract

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
  The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from. Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code. We propose that fiber bundles provide a uniform interface for describing a variety of underlying data - tables, images, networks, etc. - in a manner that independently encodes the mathematical structure of the topology and the fields of the dataset. Modeling the data structures that store the datasets as sheaves provide a method for specifying visualization methods that are designed to work regardless of how the dataset is stored - whether the data is on disk, distributed, or on demand. Specifying the visualization library components as natural transforms of sheaves means that the constraints that the component must satisfy to be structure preserving can be specified as the set of morphisms on the data and graphic sheaves, including the structure on the topology and fields of the data.
  Using category theory to formally express how visual elements are constructed means we can translate those expectations into code, which can then be used to enforce the expectation that a visualization tool is faithfully translating between numbers and charts.
\end{abstract}

@@

\begin{LaTeXdescription}
  \item[\textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we use color in definitions, equations, and visualizations to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.}] is a set of values of the same type, e.g. one column of a table or the pixels of an image
  \item[\textcolor{base}{\textbf{topology}}] is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
\end{LaTeXdescription}

@@

The function $\sigma$ is composed with $\pi$ such that $\pi^{-1}(\sigma(C)) \subseteq \mathscr{U}$; this composition induces a domain bundle $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ that associates a field name $c \in C$ with its corresponding domain $\pi^{-1}_{\sigma}(C) \subseteq \mathscr{U}_{\sigma}$. He defines a \textit{record} as a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ from column name to type domain and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$. Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$. Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from an indexing space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his construction of a table generalizes to our notion of a dataset.

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
  The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from. Motivated by wanting to make tools that are more flexible while still rigorously structure preserving, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that translates to code. We propose that fiber bundles provide a uniform interface for describing a variety of underlying data - tables, images, networks, etc. - in a manner that independently encodes the mathematical structure of the topology and the fields of the dataset. Modeling the data structures that store the datasets as sheaves provide a method for specifying visualization methods that are designed to work regardless of how the dataset is stored - whether the data is on disk, distributed, or on demand. Specifying the visualization library components as natural transforms of sheaves means that the constraints that the component must satisfy to be structure preserving can be specified as the set of morphisms on the data and graphic sheaves, including the structure on the topology and fields of the data. Using category theory to formally express how visual elements are constructed means we can translate those expectations into code, which can then be used to enforce the expectation that a visualization tool is faithfully translating between numbers and charts.
\end{abstract}

@@

\begin{LaTeXdescription}
  \item[\textcolor{fiber}{\textbf{field}}]\footnote{In this paper, color in definitions, equations, and visualizations is used to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.} is a set of values of the same type, e.g. one column of a table or the pixels of an image
  \item[\textcolor{base}{\textbf{topology}}] is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
\end{LaTeXdescription}

@@

The function $\sigma$ is composed with $\pi$ such that $\pi^{-1}(\sigma(C)) \subseteq \mathscr{U}$; this composition induces a domain bundle $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ that associates a field name $c \in C$ with its corresponding domain $\pi^{-1}_{\sigma}(C) \subseteq \mathscr{U}_{\sigma}$. He defines a \textit{record} as a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ from column name to type domain and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$\footnote{This paper follows the convention that $\Gamma$ denotes a set.}. Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$. Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from an indexing space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his construction of a table generalizes to our notion of a dataset.
